---
title: "PA1_template"
author: "Harry Werder Coursera"
date: "Tuesday, February 10, 2015"
output: md_document
---
# My Assignement 1 for "Reproducible Research"

# Part 1: What is mean total number of steps taken per day? (Without NA)
## 1.Step: Reading the data
Set the working directory which is on the Desktop and load the raw data
```{r}
setwd("C:/Users/Harry/Desktop/5- Reproducability Research/Assignment 1")
rawdata <- read.csv("activity.csv")
class_steps <- class(rawdata$steps)
class_date <- class(rawdata$date)
class_interval <-class(rawdata$interval)
numberrows <- nrow(rawdata)
```

The data has three columns and the following data types:  
steps: `r class_steps`  
date: `r class_date`  
interval: `r class_interval`  
and `r numberrows` measurements  

## 2. Step: Tidy the data up
We build now our tidy data set:
- date is transformed into a date format
- interval into a factor
- all steps with NA value are removed
```{r, echo=TRUE}
tidydata <- rawdata
tidydata$date <- as.Date(as.character(tidydata$date))
tidydata$interval<-as.factor(tidydata$interval)
tidydata <- tidydata[complete.cases(tidydata),]
nowrows <- nrow(tidydata)
```
Now our data is tidy and has `r nowrows` of measurements.

## 3. Step: What is the average daily activity pattern? 
### Analyse the data  
We will use the dplyr-package for analysis and ggplot for graphics

### 3.1. Calculate the number of steps for each day  
```{r, echo=TRUE, results="hide"}
library(dplyr)
library(ggplot2)
```
``` {r echo=TRUE}
result <- summarise(group_by(tidydata,date),sum(steps))
colnames(result) <- c("date","sum_steps")
```
Lets explore the result:
```{r, message=FALSE}
library(ggplot2)
qplot(data= result, date, sum_steps, main="Number of steps taken each day")
```

### 3.2. Lets have a look at the result with a histogram  
The historgram shows the distribution:
``` {r}
qplot(data=result, sum_steps, main="Histogram of steps taken each day")

```

### 3.3. Lets have a lock at the mean of steps and median of steps for each day
```{r}}
result2 <- summarise(result,mean(sum_steps))
result3 <- summarise(result,median(sum_steps))
Meansteps <- as.character(result2)
Mediansteps <- as.character(result3)
```
  
This is the **mean** of steps for each day:  **`r Meansteps `  steps**.  
This is the **median** of steps for each day:   **`r Mediansteps ` steps**.

## 4. What is the average daily activity pattern?

### 4.1. Eplore the activity  
On the x-axis: intervals of a day  
On the y-axis: mean of interval across all days  
```{r}
result4 <- summarise(group_by(tidydata,interval),mean(steps))
colnames(result4) <- c("interval","average_steps")
qplot(data=result4,interval,average_steps, main="Time series plot of average steps taken")
Max_average <- max(result4$average_steps)
Max_interval <- subset(result4, Max_average==average_steps)
```
The maximum average number of steps is `r Max_average `   
at **interval `r Max_interval$interval `**.

# Part 2: Imputing missing values
We start with the raw data again and make them clean.  
## 1.Step: Cleaning and counting NA rows
``` {r}
tidydata2 <- rawdata
tidydata2$date <- as.Date(as.character(tidydata2$date))
tidydata2$interval<-as.factor(tidydata2$interval)
nowrows2 <- nrow(tidydata2)
NAinData <- nowrows2 - nowrows
inpercent <- round(NAinData/nowrows2*100,0)
```
The data set has `r NAinData` rows with NA data: this is `r inpercent`% of all rows.

## 2. Step: Replace strategy  
Let us use a **strategy to replace the NA values** with the **mean value** for the interval across all days.  
```{r}
result5 <- summarise(group_by(tidydata2,interval),mean(steps, na.rm=TRUE))
colnames(result5) <- c("interval","average_steps")
#select the NA data
bad<- is.na(tidydata2$steps)
allNA<-subset(tidydata2, bad)
#replace the NA with the mean for this interval
for (i in 1:nrow(allNA)){
    b <- allNA[i,3]
    a <- subset(result5,b==result5$interval)
    newsteps <-round(a$average_steps,0)
    allNA[i,1] <- newsteps
}
#new data with NA relaced by mean: take the tidydata with NA and add our allNA with the means
tidydata3 <- rbind(tidydata,allNA)
```
## 3. Analyse the data with replaced NA by mean of interval

### 3.1. Histrogram
The historgram shows the distribution; number of steps by day
``` {r}
result6 <- summarise(group_by(tidydata3,date),sum(steps))
colnames(result6) <- c("date","sum_steps")
qplot(data=result6, sum_steps, main="Histogram of total number of steps taken each day")
```

### 3.2. Lets have a lock at the mean of steps and median of steps for each day
```{r}
result21 <- summarise(result6,mean(sum_steps))
result31 <- summarise(result6,median(sum_steps))
Meansteps1 <- as.character(result21)
Mediansteps1 <- as.character(result31)
```
  
This is the mean of steps for each day:  `r Meansteps1 `  steps.  
This is the median of steps for each day:   `r Mediansteps1 ` steps.

# Compare result with "NA removed"" and with "NA replaced by mean of interval"
##4.1 Mean-Median per day
Let's build a table for comparision for mean and median:
```{r}
data.frame(Cases=c("NA removed", "NA replaced by mean"), Number_of_cases= c(nowrows,nowrows2), Mean_value= c(Meansteps,Meansteps1), Median_value= c(Mediansteps, Mediansteps1))

```

Conclusion: there is only a slight effect on relpacing NA values with the mean of the interval.

#4.2. Total number of steps per day
We take the data set without NA and calculate the total number of steps per day.
We then take the data set with replaced NA by mean and calculate the total number of steps per day.
```{r}
stepsNAremoved <- summarise(group_by(tidydata,date),sum(steps))
colnames(stepsNAremoved)<-c("dates","steps")
stepsNAreplaced <- summarise(group_by(tidydata3,date), sum(steps))
colnames(stepsNAreplaced)<-c("dates","steps_NA_corrected")
result7<- merge(stepsNAremoved, stepsNAreplaced, by="dates",all=TRUE)
result7$impact <- result7$steps_NA_corrected-result7$steps
```

The replacement of missing values (NA) by a mean values has given days without measurement now a value.  
```{r}
result7
```


# Part 3: Are there differences in activity patterns between weekdays and weekends?  

We use for this analysis the data set with the replaced NA by the mean of the interval (of all days).

My Question: is there a difference in weedays (Mo-Fr) to weekend (Saturday & Sunday)?

```{r}
tidydata3$day <- weekdays(tidydata3$date,abbreviate = FALSE)
# decide if weekday or weekend
tidydata3$day["Samstag"==tidydata3$day]<-"weekend"
tidydata3$day["Sonntag"==tidydata3$day]<-"weekend"
tidydata3$day["weekend"!=tidydata3$day]<-"weekday"
tidydata3$day <- as.factor(tidydata3$day)
# create the summary for the plot
result8<-summarise(group_by(tidydata3,day,interval),mean(steps))
colnames(result8) <- c("day","interval","sum_steps")  #adjust the names of columns

qplot(data=result8,as.numeric(interval),sum_steps, col=day, geom="line", main= "Panel Plot")+facet_grid(day~.)+xlab("interval of the day")+ylab("average of steps per interval")+guides(col=FALSE)
